Proposed Solution

### 2016-11-10 Ennan-Proposal/solutions.tex
autocommit

\subsection{Derivation of Probabilistic Rules}
\label{subsec-rules}

@@

\para{A strawman solution.}
We first present a strawman solution (employed by 
our existing work ConfigC~\cite{santolucitoCAV}) that uses 
a set of correct configuration files as a learning set, 
from which it is possible to derive rules 
that must hold with absolute certainty. 
%However, as mentioned in Sec.~\ref{sec:prelim}, 
%it is difficult to obtain a set of files 
%that is both guaranteed to be without errors
%and large enough to learn sufficiently many rules.
%This usually requires manual verification of the learning set, 
%which is prone to error.
Under such an assumption, the strawman solution
only considers a rule if it holds over exactly every file in 
the learning set. This behavior can be formally described as follows:
\begin{small}
\begin{align*}
C :=&\ \text{Correct Learning Set}\\
LR :=&\ \text{Learned Rules} :: \{\textrm{Rule}\}\\
RR :=&\ \text{Reported Rules} :: \{\textrm{Rule}\}\\ 
C =&\ \text{\{Configuration Files in Intermediate Representation\}}\\
LR =&\ \{ r\ \mid \forall file \in C,\ holds(r,file)\} \\
RR =&\ \{ r\ \mid r \in LR \ \land \neg\ holds(r,\textrm{user file}) \}\\
\end{align*}
\end{small}

where, the $:: \{Rule\}$ indicates that $LR$ (\ie, the learned rule set) 
and $RR$ (\ie, the reported rule set) are sets of rules.
A rule is only committed to the learned rule set, $LR$, if 
the rule holds on all files in the training set $C$, 
as denoted by the $\forall$ quantifier. 
A broken rule is reported to the user in the $RR$ rule set if the rule is in $LR$, 
but does not hold on the input user file.

In general, each rule can be viewed as a mapping from a tuple of 
entry keys $(k_1, \ldots, k_n)$ in the intermediate representation of a 
configuration file to a probability function, $P_{rel}$. 
This can be represented by a rule
$r := (k_1, \ldots, k_n) \rightarrow P_{rel}$.
$P_{rel}$ is a probability function specific to the relation to assess.
The problem with the strawman solution is that the probability functions are restricted to be 
boolean, which is to say that the relation can either be unsatisfied or satisfied, either 0 or 1. 
For instance, for the rule that the value of {\tt max\_connections} must 
be higher than the value of {\tt mysql.max\_persistent}, 
the rule is denoted by
({\tt max\_connections}, {\tt mysql.max\_persistent}) $\to 1$, where 
$B_{>}$ is set to 1. In our approach (see below), 
we permit the probability functions to take on any value from 0 to 1.

\para{Probabilistic approach.}
In \app's learner module, we release the assumption in the above
strawman solution, which means the rule learning mechanism is tolerant 
enough to accept a dataset of possibly incorrect configuration files. 
Rather than manually correcting each file, 
we extend the previous formalism to run probabilistic learning
on our intermediate representations (generated by the translator).

\begin{algorithm}[t]
\caption{Probabilistically Learn Rules: \textbf{learnRules()}}
\label{alg:plearn}
\begin{footnotesize}
  \KwIn{$\theta$ is the list of possible types.}
  \KwIn{$U$ is the list of possible untyped specifications.}
  \KwIn{$C$ is a set containing intermediate representations
     of configuration files}

  Let $R = \{\}$ be {\em rule set}\;
  Let $\Pi$ be {\em probabilities associated with rule set}\;
  \For{each file $F$ in $C$}{
      Initialize $opt$ flag\;
      \For{each $z \in \theta \cup U$}{
          $Q$ contains candidate entries to consider\;
          $RM$ contains entries in remove set\;
          \If{$z$ is a type}{
              $opt = 1$\; 
              $Q = filter(F,\tau)$\; /* {\em Q contains $F$ entries with
$\tau$ in type list} */\\
          }\ElseIf{$z$ is an untyped spec}{
              $opt = 0$\;
              $Q = F$\;
          }
          %\EndIf
          \For{each entry $e$ in $Q$}{
              \If{$g(e)$ is false}{
                  continue to next entry\;
              }\Else{
                  $CR = \textrm{constructRules}(e, z, opt, Q \backslash
RM)$\;
                  $R = R  \cup CR$\;
                  For each $r \in CR$, update $\Pi$\;
              }
              $RM = RM \cup e$\;
          }
        }
  }
  return $(\Pi,R)$\;
\end{footnotesize}
\end{algorithm}

\com{
\begin{algorithm}[t]
\caption{Construct Rules: constructRules($e, z, opt, S$)}
\label{alg:crules}
\begin{footnotesize} 
    \KwIn{entry $e$ from intermediate representation.}
    \KwIn{type or untyped specification $z$}
    \KwIn{option $opt$, set to 1 if template or 0 if untyped spec}
    \KwIn{set of entries $S$}

    $R = \{\}$\;
    Set $j$ as an iteration variable\;
    \If{opt = 1}{
        run the following sequence \textbf{L} over 
        all elements of $templates(z)$\;
    }\ElseIf{opt = 0}{
        $j = z$\;
    }
    Beginning of \textbf{L}: 
     \State Let $n$ be the arity of $j$.
      \State Let $P$ be the set of permutations of distinct entries from $S$, of size $n-1$.
      \For{each $p \in P$}
        \State Let $p$ be represented by $(e_1, e_2, \ldots, e_{n-1})$.
         \For{$i = 1:n$}
         \State Create a $n$-tuple by placing $e$ at the $i$-th position and filling the remaining entries with $p$, in order. Let this $n$-tuple be $m = (e_1, e_2, \ldots, e_i, \ldots, e_{n-1})$.
         \State $r = \textrm{constructRule}(m)$
         \If{$r$ is satisfied}
                $R = R \cup r$
         \EndIf
        \EndFor
     \EndFor
     \State End of \textbf{L}. 
     \State return $R$
  \end{algorithmic}
\end{footnotesize}
\end{algorithm}
}

Our probabilistic approaches for learning missing entry, 
ordering, and fine-grained value correlation rules stem 
from existing work with building 
non-probabilistic versions of these rule-learning algorithms. 
For each of these rules, 
we consider all possible permutations of keys that appear in every 
file which are appropriately typed, and for our learning process, calculate the likelihood that each of 
these permutations constitute a rule. 
Rule patterns in the example set that appear frequently are accepted,
so they can be used to evaluate new files.

More formally, this approach can be defined as follows:

\begin{small}
\begin{flalign*}
I\ \ :=&\ \text{Incorrect Learning Set}\\
LP :=&\ \text{Learned Probabilistic Rules :: \{P\_Rule\}}\\
RP :=&\ \text{Reported Probabilistic Rules :: \{P\_Rule\}}\\
I =&\  \text{\{Configuration Files in Intermediate Representation\}}\\
(\Pi, LP) =&\ \text{learnRules}(I)\\
RP =&\ \{ r\ \mid r \in LP\ \land \Pi(r)>p \land \neg holds(r, userfile) \}
\end{flalign*}
\end{small}

Note that the formalism relies on Algorithm~\ref{alg:plearn}. The algorithm considers both typed templates and
untyped specifications. As a similar iteration logic is used for both, an $opt$ flag distinguishes between the two situations.
constructRule in Algorithm~\ref{alg:crules} refers to the creation of a rule from a tuple of entries, 
$m = (e_1, e_2, \ldots, e_i, \ldots, e_{n-1})$.
The rule creation is formally represented as $m(k(e_1), k(e_2), \ldots, k(e_i), \ldots, k(e_{n-1}))$. 
In order to determine if the rule satisfied, the relation associated with $m(v(e_1), v(e_2), \ldots, v(e_i), \ldots, v(e_{n-1}))$ is evaluated for truth.
The rule can be template-generated, for instance {\tt max\_connections} > {\tt mysql.max\_persistent}, or
it can be untyped-specification-generated, for instance the ordering that {\tt recode.so} must come before 
{\tt mysql.so}.

@@

A rule $r$ will be reported as broken if the probability the rule is
correct, $\Pi(r)$, is greater than some user defined constant, $p$. This
constant can be adjust to the user's preference. A small $p$ will
increase the likelihood of finding an error, but also increase the
number of false positives that are reported.

>>>>

\subsection{Derivation of ConfigC Rules}
\label{subsec-rules}

@@

%However, as mentioned in Sec.~\ref{sec:prelim}, 
%it is difficult to obtain a set of files 
%that is both guaranteed to be without errors
%and large enough to learn sufficiently many rules.
%This usually requires manual verification of the learning set, 
%which is prone to error.
This only considers a rule if it holds over exactly every file in 
the learning set. This behavior can be formally described in Fig~\ref{fig:configC}.

\begin{figure}[!h]
\begin{small}
\belowdisplayskip=-15pt
\abovedisplayskip=-2pt
\begin{align*}
C :=&\ \text{Correct Learning Set}\\
C =&\ \text{\{Configuration Files in Intermediate Representation\}}\\
LR :=&\ \text{Learned Rules} :: \{\textrm{Rule}\}\\
LR =&\ \{ r\ \mid \forall file \in C,\ holds(r,file) \\
&\qquad \ \exists file \in C,\ nontriviallyHolds(r,file)\} 
\end{align*}
\end{small}
\caption{The definition of which rules should be learned by ConfigC}
\label{fig:configC}
\end{figure}

where, the $:: \{Rule\}$ indicates that $LR$ (\ie, the learned rule set) 
is a sets of rules.
The predicate $holds$ indicates that the rule should always be true -
  for instance {\tt max\_connections} $>$ {\tt max\_persistent}.
The predicate $nontriviallyHolds$ indicates that the rule should be true in a meaningful way.
For instance, a file may only contain the keyword {\tt max\_connections} - the previous rule then holds vacuously,
 because we say {\tt max\_connections} is always greater than {\tt max\_persistent} if no value is provided for {\tt max\_persistent} in that file.

\subsection{Derivation of ConfigV Rules}
In \app's learner module, we release the assumption in the above
solution, which means the rule learning mechanism is tolerant 
enough to accept a dataset of possibly incorrect configuration files. 
Rather than manually correcting each file, 
we extend the previous formalism to run probabilistic learning
on our intermediate representations (generated by the translator). 
Our probabilistic approaches for learning missing entry, 
ordering, and fine-grained value correlation rules stem 
from existing work with building 
non-probabilistic versions of these rule-learning algorithms. 
For each of these rules, 
we consider all possible permutations of keys that appear in every 
file which are appropriately typed, and for our learning process, calculate the likelihood that each of 
these permutations constitute a rule. 
Rule patterns in the example set that appear frequently are accepted,
so they can be used to evaluate new files.

\begin{figure}[!h]
\begin{small}
\belowdisplayskip=-15pt
\abovedisplayskip=-2pt
\begin{align*}
U\ :=&\ \text{Unlabeled Learning Set}\\
U =&\ \text{\{Configuration Files in Intermediate Representation\}}\\
LP :=&\ \text{Learned Probabilistic Rules :: \{P\_Rule\}}\\
LR =&\ \{ r\ \mid \Pi file \in C,\ holds(r,file)\ \land \\
&\qquad  \ \Psi file \in C,\ nontriviallyHolds(r,file)\} 
\end{align*}
\end{small}
\caption{The definition of which rules should be learned by ConfigV when learning over a set of possibly correct, and possibly incorrect files.}
\label{fig:configV}
\end{figure}

We use $\Pi$ and $\Psi$ in Fig~\ref{fig:configV} to indicate that the rule holds for some critical portion of the files.
The inuition behind these quantifiers are that we want to consider rules which are true in majority of cases ($\Pi$), 
and for which we have at least some examples for which that rule is nontrivially true ($\Psi$).
The exact definition of $\Pi$ and $\Psi$ is a research question still to be solved.
What exactly is the cutoff value that will yield the best results in learning.
We plan to try multiple strategies to find these values.
One approach may be to use machine learning for program optimization, with a tool like OpenTuner~\cite{ansel:pact:2014}.
It is also possible that this value differs depending on the quality of the learning set and/or the particular type of rule being learned.

\iffalse
Note that the formalism relies on Algorithm~\ref{alg:plearn}. The algorithm considers both typed templates and
untyped specifications. As a similar iteration logic is used for both, an $opt$ flag distinguishes between the two situations.
constructRule in Algorithm~\ref{alg:crules} refers to the creation of a rule from a tuple of entries, 
$m = (e_1, e_2, \ldots, e_i, \ldots, e_{n-1})$.
The rule creation is formally represented as $m(k(e_1), k(e_2), \ldots, k(e_i), \ldots, k(e_{n-1}))$. 
In order to determine if the rule satisfied, the relation associated with $m(v(e_1), v(e_2), \ldots, v(e_i), \ldots, v(e_{n-1}))$ is evaluated for truth.
The rule can be template-generated, for instance {\tt max\_connections} > {\tt mysql.max\_persistent}, or
it can be untyped-specification-generated, for instance the ordering that {\tt recode.so} must come before 
{\tt mysql.so}.

@@

A rule $r$ will be reported as broken if the probability the rule is
correct, $\Pi(r)$, is greater than some user defined constant, $p$. This
constant can be adjust to the user's preference. A small $p$ will
increase the likelihood of finding an error, but also increase the
number of false positives that are reported.
\fi
